---
title: 'Lab 3: Mark-Recapture Introduction'
author: "Put your name here"
output:
  pdf_document: default
  word_document: default
---

Turn in via Canvas next Friday by the end of the day. Submit either a knitted pdf, html, or doc file. Knit frequently as you go. It's easier to trouble shoot along the way rather than when you have lots of untested code.

**Do not print full dataframes unless I ask for them or I will not read your homework**

Your instructor will provide instruction about R-projects and working directories.

```{r, eval=FALSE}
getwd()

setwd("C:/Users/djhocking/Documents/Quant_Vert_Pop/")

```


For this lab we will be using turtle Mark-Recapture data from the C&O Canal in Cumberland, MD collected by Frostburg State University students. The data are collected using a robust design spatial capture-recapture approach with turtles receving individual marks. 

Your instructor will provide more details about the study.

We will use these data for a variety of mark-recapture labs to examine the details and inference we get from different study designs. This week we will pretend that the data were collected using a closed population design. In later labs we will allow for population change explicitly and more complicated individual capture history information.

First, we will read in the full turtle data from the first three primary session (summer 2018, fall 2018, spring 2019). Then we will summarize the data as if we just batch marked the turtles:

```{r, message=FALSE, warning=FALSE}
library(dplyr)
turtles <- read.csv("Data/turtle_scr.csv", stringsAsFactors = FALSE)
# str(turtles)

turtle_lp <- turtles %>%
  group_by(species, primary, secondary, batch_recap) %>%
  summarise(count = n())

# turtle_lp
```

## 1. Use the data above to fill out this table for Painted Turtles (`CPIC`):

| PP  | Day  | $C_t$  | $R_t$  | New Marks  | $M_t$  |
|---|---|---|---|---|---|
| 1  | 1  |   |   |   |   |
| 1  | 2 |   |   |   |   |
| 1  | 3 |   |   |   |   |
| 1  | 4 |   |   |   |   |
| 2  | 1 |   |   |   |   |
| 2  | 2 |   |   |   |   |
| 3  | 1 |   |   |   |   |
| 3  | 2 |   |   |   |   |
| 3  | 3 |   |   |   |   |

## 2. Use these data to estimate the population size ($\hat N$) in each primary period with the Lincoln-Petersen Estimator. Just use the first 2 days of each primary period, pretending that we did a traditional Lincoln-Petersen design with just two sampling days (secondary sessions).

## 3. Now use these data to estimate $\hat N$ using the Chapman Method for each primary period. Again only use the first two secondary sessions (days) for the Chapman method. 

## 4. Finally, use these data to estimate $\hat N$ with the Schnabel Method for each primary period. For this method you will use all secondary sessions within a primary period.

## 5. Make a plot of estimated abundances (y-axis) over time (x-axis = primary period) with one line for each method.


```{r}
abund_df <- read.csv("Data/lab3.csv")

library(ggplot2)

ggplot(data = abund_df, aes(PP, N, color = method)) + geom_point() + geom_line()

```

```{r}
df <- read.csv("Data/lab3_wide.csv")

ggplot(data = df, aes(PP, LP)) + geom_point() + geom_line() +
  geom_point(aes(PP, Chapman), color = "blue")

```

## 6. Compare and contrast the estimates from each method and describe the pros and cons of each and differences in inference from the different methods. (Discuss your results).

## 7. Write out the probability for this capture history assuming that $p(.) = c(.)$ for a closed population (see page 98-99 in the textbook along with class notes):

00101010

## 8. Write out the probability for this capture history assuming that $p(.) \neq c(.)$ for a closed population:

00101010

## 9. Write out the probability for this capture history assuming that $p(t) = c(t)$ for a closed population:

00101010

## 10. Write out the probability for this capture history assuming that $p(t) \neq c(t)$ for a closed population:

00101010


## 11. Using the simple_closed1.inp data, run a closed population Huggins model in program MARK assumming p(.) = c(.). Display the results on the true scale.

## 12. Using the simple_closed1.inp data, run a closed population Huggins model in program MARK assumming p(.), c(.). Display the results on the true scale.

Delete this example after viewing how to set up the table after pasted in from program MARK:


  Parameter                  Estimate       Standard Error      Lower           Upper
 --------------------------  --------------  --------------  --------------  --------------
     1:p                      0.5195804       0.0093423       0.5012514       0.5378568                          
     2:p                      0.5487627       0.0134246       0.5223385       0.5749148                          
     3:p                      0.6112901       0.0195768       0.5723037       0.6489030                          
     4:p                      0.5684643       0.0319044       0.5051694       0.6295993                          
     5:p                      0.4038452       0.0481135       0.3140690       0.5005568                          
     6:p                      0.3225809       0.0593673       0.2185014       0.4478278                          
     7:p                      0.4285700       0.0763621       0.2892890       0.5801682                          
     8:p                      0.4166702       0.1006373       0.2408554       0.6165854                          
     9:p                      0.2142841       0.1096602       0.0707071       0.4943233                          
    10:p                      0.4545381       0.1501408       0.2027383       0.7319552                          
    11:p                      0.4999754       0.2040980       0.1679499       0.8320226                          
    12:p                      0.3333344       0.2721765       0.0433664       0.8465053                          
    13:p                      0.9999444       0.0126608      0.2956149E-189   1.0000000                          
    14:p                      0.0657492       5.2975965       0.2727215E-74   1.0000000                          
    15:p                      0.0657491       4.8740591       0.2017527E-68   1.0000000                          
    16:c                      0.4885599       0.0129672       0.4631955       0.5139833                          
    17:c                      0.5183034       0.0105574       0.4975914       0.5389526                          
    18:c                      0.3417335       0.0092678       0.3238074       0.3601234                          
    19:c                      0.3457910       0.0090599       0.3282548       0.3637567                          
    20:c                      0.3498928       0.0090165       0.3324329       0.3677645                          
    21:c                      0.2274663       0.0078967       0.2123611       0.2433140                          
    22:c                      0.2383639       0.0080009       0.2230379       0.2543981                          
    23:c                      0.2456078       0.0080687       0.2301385       0.2617634                          
    24:c                      0.1705861       0.0070471       0.1572144       0.1848457                          
    25:c                      0.1706378       0.0070418       0.1572757       0.1848860                          
    26:c                      0.1753587       0.0071144       0.1618478       0.1897422                          
    27:c                      0.2197340       0.0077453       0.2049303       0.2352907                          
    28:c                      0.2276224       0.0078404       0.2126218       0.2433541                          
    29:c                      0.2346154       0.0079238       0.2194424       0.2505008                          
       N                      2860.0001       0.0243717       2860.0000       2860.0656   


## 13. Using the simple_closed1.inp data, run a closed population Huggins model in program MARK assumming p(t) = c(t). Display the results on the true scale.

## 14. Using the simple_closed1.inp data, run a closed population Huggins model in program MARK assumming p(t), c(t). Display the results on the true scale.

## 15. Thoroughly describe and interpret the results of numbers 11-14. Compare and contrast the results and the results of the different models and do multimodel comparison using AIC.

