---
title: "Lab 4 Tutorial: Hierarchical Models"
author: "Daniel Hocking"
output: pdf_document
---

## Background

**Read Chapters 14, 15, 17, 18 in Powell and Gale**

## Key Terms

* Census
* Survey
* Count
* Abundance (estimate)
* Catch-per-unit-effort (CPUE)
* Index of abundance
* Detection probability
* Availability probability

## Methods

* Distance sampling
* Double-observer
* Removal sampling (also called depletion sampling)
* Repeated counts (*N*-mixture model)
* Occupancy modeling

## Occupancy Model

Repeated presence-absence data to account for imperfect detection

### Example

*formula in `unmarked`*
Double right-hand side formula describing covariates of detection and occupancy in that order.

```{r}
# --------------------- Analysis of real data ----------------------------
# This a subset of point-count data from Chandler et al. (Auk 2009)
# alfl is Alder Flycatcher (Empidonax alnorum)
# PART I. Import and format data
alfl.data <- read.csv("Data/alfl05.csv", row.names=1)
str(alfl.data)
# Pull out count matrix and covert to binary
alfl.y <- alfl.data[,c("alfl1", "alfl2", "alfl3")]
alfl.y1 <- alfl.y # Make a copy
alfl.y1[alfl.y>1] <- 1

# Standardize site-covariates
woody.mean <- mean(alfl.data$woody)
woody.sd <- sd(alfl.data$woody)
woody.z <- (alfl.data$woody-woody.mean)/woody.sd
struct.mean <- mean(alfl.data$struct)
struct.sd <- sd(alfl.data$struct)
struct.z <- (alfl.data$struct-struct.mean)/struct.sd

# Create unmarkedFrame
library(unmarked)
alfl.umf <- unmarkedFrameOccu(y=alfl.y1,
    siteCovs=data.frame(woody=woody.z, struct=struct.z),
    obsCovs=list(time=alfl.data[,c("time.1", "time.2", "time.3")],
                 date=alfl.data[,c("date.1", "date.2", "date.3")]))
# Here’s an easy way to standardize covariates after making the UMF
obsCovs(alfl.umf) <- scale(obsCovs(alfl.umf))
summary(alfl.umf)

# PART II. Model fitting
(fm1 <- occu(~1 ~1, alfl.umf))
backTransform(fm1, type="state")
backTransform(fm1, type="det")
(fm2 <- occu(~date ~woody, alfl.umf))

```

```{r}
backTransform(linearComb(fm2, coefficients = c(1,-2), type = "state"))

(exp(1.5587 + 2.0502 * -2)) / (1 + exp(1.5587 + 2.0502 * -2))

```


## Distance Sampling

### Example

```{r}
library(unmarked)

# Load and format the data
data(issj) # island scrub jay
head(issj)
covs <- issj[,c("elevation", "forest", "chaparral")]
area <- pi*300^2 / 10000 # just converts distance in meters to hectares
jayumf <- unmarkedFrameDS(y=as.matrix(issj[,1:3]),
                          siteCovs=data.frame(covs, area),
                          dist.breaks=c(0,100,200,300),
                          unitsIn="m", survey="point")
# Fit model
fm1 <- distsamp(~chaparral ~chaparral + offset(log(area)),
                jayumf, keyfun="halfnorm", output="abund")
fm1
```


## Abundance with an *N*-mixture model

### Example

```{r}
# PART 1: Set-up the data for analysis
#
#
# -------------------------- Format data ---------------------------------
# This a subset of point-count data from Chandler et al. (Auk 2009)
# alfl is Alder Flycatcher (Empidonax alnorum)

# Import data and check structure
#alfl.data <- read.csv("alfl05.csv", row.names=1)
alfl.data <- read.csv("Data/alfl05.csv", row.names=1)

str(alfl.data)

# Pull out count matrix 
# No need to covert to binary as we did for occupancy model

alfl.y <- alfl.data[,c("alfl1", "alfl2", "alfl3")]

# Standardize site-covariates
woody.mean <- mean(alfl.data$woody)
woody.sd <- sd(alfl.data$woody)
woody.z <- (alfl.data$woody-woody.mean)/woody.sd

struct.mean <- mean(alfl.data$struct)
struct.sd <- sd(alfl.data$struct)
struct.z <- (alfl.data$struct-struct.mean)/struct.sd


# Create unmarkedFrame
library(unmarked)
alfl.umf <- unmarkedFramePCount(y=alfl.y,
    siteCovs=data.frame(woody=woody.z, struct=struct.z),
    obsCovs=list(time=alfl.data[,c("time.1", "time.2", "time.3")],
                 date=alfl.data[,c("date.1", "date.2", "date.3")]))
summary(alfl.umf)


# Here's an easy way to standardize covariates after making the UMF
obsCovs(alfl.umf) <- scale(obsCovs(alfl.umf))
summary(alfl.umf)


#
#
# PART 2: Fit some models
#
# -------------------------- Model fitting  -----------------------------

(fm1 <-  pcount(~1 ~1, alfl.umf))
backTransform(fm1, type="state")
backTransform(fm1, type="det")

# model with date affecting detection and woody vegetation affecting abundance
(fm4 <- pcount(~date ~woody, alfl.umf))
```

## Removal Sampling

### Example

We can use data that come loaded with a package using the `data()` function. You can get a list of all the data available with currently loaded packages in R using `names(data())`. - not working!

For example if we load `unmarked` and look at the available data

```{r, message=FALSE, warning=FALSE}
library(unmarked)
names(data())
```

```{r}
library(unmarked)
data(issj)
str(issj)

help("issj")

data(ovendata)
head(ovendata.list[[1]])
head(ovendata.list[[2]])
```




