---
title: "Homework 2: Linear Regression"
author: "Your_Name_Here"
output: pdf_document
---

---
title: "Homework 2: Plotting & Linear Regression"
author: "BIOL 414/514"
date: "Quantitative Analysis of Vertebrate Populations"
output:
  tufte::tufte_handout:
    df_print: kable
  tufte::tufte_html: default
urlcolor: blue
bibliography: bibliography.bib
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo = FALSE}
library(tufte)
```

Turn in via Canvas next Friday by the end of the day. Submit both the .Rmd file and either a knitted pdf, html, or doc file.

## 1. Using the `iris` data, make a new dataframe called `iris_petal` with just the petal length and width columns.

```{r}
str(iris)
```

Based on the structure of iris I see that there are 150 observations of 5 variables. 

## 2. Now add a new column to `iris_petal` called `Random` filled with normally distributed random numbers.

## 3. Display the first 15 rows of the `iris_petal` dataframe.

## 4. Print a summary of the `iris_petal` dataframe.

## 5. Read the `mallard_counts.csv` file into R and save the dataframe as an object call `ducks`. Be sure to use `stringsAsFactors = FALSE`. 

## 6. Print at the structure of `ducks`. How many rows and columns does it have? 

## 7. Plot the elevation vs. counts and include a linear regression line using ggplot

## 8. Run a linear regression of the effects of elevation and forest on count and print the summary table.

## 9. Are the linear model assumptions met and how do you know (Hint: make diagnostic plots using `plot(Model_Name)` and interpret the output. You won't lose points if you correctly explain the plots even if I disagree with your judgement on whether things are normal or homogenous enough.

## 10. What is the effect of elevation on the number of ducks observed? What's the effect of forest? Are they statistically significant and how do you know?



