---
title: "Lab 3: Generalized Linear Models"
author: "Your_Name_Here"
output: pdf_document
---

Turn in via Canvas next Friday by the end of the day. Submit either a knitted pdf, html, or doc file (you can submit the .Rmd file if knitting doesn't work).

1. Import (read in via a funtion in R, don't use the import funtion in RStudio) the `hares.txt` data. Look at (print) the structure of the data.

```{r}

```

Get rid of missing independent data with the following code (I used `hares` as my dataframe when I read the data in but you can switch it out if you used something else).

```{r}
hares <- hares[which(!is.na(hares$count1)), ] # this takes all the columns for the rows where count1 is not missing (NA)
```

2. Run a linear model with a normal distribution to test the effect of year, elevation, landuse, and region on the count (count1) of bunnies.

3. Does this model fit these data? Why or why not? How do you know? Include any tables or visualization you think appropriate.

4. Examine the residual plots and consider the assumptions of the model. What's wrong with this model? What is a better approach and why?

5. Run a linear model with a normal distribution to test the effect of year, elevation, landuse, and region on the **mean.density** of bunnies.

6. How does this model fit the data?

7. Run a GLM with a Poisson distribution on the count1 data as a function of year, elevation, landuse, and region.

8. What are the effect sizes (coefficient estimates and their uncertainties (SE))? Do you believe the model (look at the residuals)?

9. Add in the effect of area? Does that make a difference? What is the affect of area? 

**Bonus**: See if you can add in the affect of area as an `offset` instead of a linear predictor. Is that any different?

10. Run a logistic regression on the presence data with the same independent data as #7. 

To get the data into presence/absence instead of counts, start with this code:

```{r}
# I called my data "hares" but you would use whatever dataframe you read the data into.
hares$presence <- ifelse(hares$count1 > 0, 1, 0)
```

11. Read in the seals data, explore the data, and make a model of the data (use count as the indendent data). Describe what dependent variables you chose, why, and what model you used and why you think it might be appropriate.



