---
title: "Homework 3: Linear Mixed Models"
author: "Your_Name_Here"
output: pdf_document
---

Turn in via blackboard next Friday by the end of the day. please title the document **hw3_yourlastname_yourfirstname**. Ideally it will be a PDF file generated from knitting an RMarkdown document.

In this lab you will be working with the RIKZ data (Zuur et al. 2007). They are data on the macrofauna species richness from five sites at each of nine beaches. The authors were interested in how  species richness is affected by exposure and NAP (height of sampling station relative to mean tidal level). However, they likely have to account for the potential correlation of data from sites collected at the same beach.


## 1. Read in the RIKZ data from blackboard using `read.table` since it's a text file and not a csv. Assign it to an object and examine the structure of the object. How many rows and columns are there?

```{r}
df <- read.table(file = "Data/RIKZ.txt", header = TRUE, stringsAsFactors = FALSE)
str(df)
ncol(df)
nrow(df)
```

45 rows and 5 columns


## 2. Use a linear regression to examine the effects of NAP on species richness. 

```{r}
lm1 <- lm(Richness ~ NAP, data = df)
summary(lm1)
```

## 3. Check if the model assumptions fit the data. Describe the results.

```{r}
plot(lm1)
```

Heterogeneity of residuals with increasing variance with larger values. Also some potential patterning in the residuals. Significant violations of normality based on the QQplot. I would not use this model with these data.

## 4. Analyze the same patterns, this time using a linear mixed effects model with a random intercept for beach. Check the model assumptions including homogenity by group (beach).

```{r}
library(lme4)
lme1 <- lmer(Richness ~ 1 + NAP + (1 | Beach), data = df)
# par(mfrow = c(1,2))
plot(lme1)

boxplot(resid(lme1) ~ df$Beach)
# par(mfrow = c(1,1))
```

## 5. Run a third model using a random intercept and random effect of beach on NAP (random slope). Check the model assumptions including homogenity by group (beach).

```{r}
lme2 <- lmer(Richness ~ 1 + NAP + (1 + NAP | Beach), data = df)
plot(lme2)
boxplot(resid(lme2) ~ df$Beach)
```

## 6. Which model would you use and why?

I would use the random intercepts and slopes model of these options but there still might be some heterogeneity of variance in the residuals and I would explore other models for these data.

## 7. Describe the results of this model.

```{r}
summary(lme2)
```

Richness decreases with increasing NAP but this affect varies modestly among beaches based on the random slope SD. There is large variation in richness among beaches (SD = 3.5) not explained by the deterministic part of the model (NAP).

 



While diving off Pūpūkea Beach you notice an algae that you have never seen before. You wonder if the parrotfish you see at other beaches are able to keep down this algal invader. You want to test the question: Do parrotfish have an effect on invasive Macroalgae cover? You suspect there might be an effect of water depth. Furthermore, any given location has an amount of native algae.  To do so, you decide to do a cage experiment, excluding parrotfish from certain locations. You have permits to work at 5 beaches, at both shallow and deep depths. You have 160 cages. How do you design this experiment?


Once we have discussed the design of the experiment, open the .csv file named “AlgaeProject.csv”. Your assignment (due next Friday, as an .R document) is to analyze the data with the appropriate statistical model and address the research question above.


