---
title: "Homework 4: Mixed Models and Data Exploration"
author: "Your_Name_Here"
output: pdf_document
---

Turn in via blackboard next Monday by the end of the day. please title the document **hw4_yourlastname_yourfirstname**. Ideally it will be a PDF file generated from knitting an RMarkdown document.

## 1. The package `tidyr` can also be used to reshape and organize data. Load the `tidyr` package. Look up the help menu for the functions `spread()` and `gather()`. The help documentation isn't very good, so doing a web search for a tutorial is likely the easiest option. Try these functions to go back and forth between long and wide format using the data `d` as we did with reshape during lab.

```{r}
library(tidyr)

```

## 2. Install and load the package `lubridate`. There is a great tutorial at [https://rpubs.com/davoodastaraky/lubridate](https://rpubs.com/davoodastaraky/lubridate). Use it to get today's date. Convert it from year-month-day format to the more familiar US format of month-day-year.


## 3. Use any dataset you want and make one histogram and one boxplot and print them side by side (hint: use the `par(mfrow = )` function.)


## 4. Use a linear mixed model to examine the effects of NAP on species richness with a random intercept for beach (from homework 3). Now use the regression equation and independent data along with the model coefficients, beach-specific random effects, and residuals to exactly recreate the original dependent data.

```{r}
df <- read.table(file = "Data/RIKZ.txt", header = TRUE, stringsAsFactors = FALSE)

lme1 <- lmer(Richness ~ 1 + NAP + (1 | Beach), data = df)
lme1_sum <- summary(lme1)
# str(lme1_sum)
```

```{r}
int_mean <- fixef(lme1)[1] # mu_alpha or lme2_sum$coefficients[1,1]
slope <- fixef(lme1)[2] # beta
int_rands_0 <- ranef(lme1)$Beach[1] # get random intercept
lme1_resids <- residuals(lme1) # get residuals
int_rands <- int_rands_0 + int_mean # get specific intercept for each beach
df_rands <- data.frame(Beach = 1:9, alpha = int_rands) # make into a dataframe with beaches labeled 
names(df_rands) <- c("Beach", "alpha_j") # rename columns


df <- merge(x = df, y = df_rands, by = "Beach", all.x = TRUE) # put specific beach intercept effect into the dataframe
df$B <- slope # add slope to the dataframe
df$resids <- lme1_resids # add residuals to the dataframe

# Use equation to calculate expected values
df$exp_y <- df$alpha_j + df$B * df$NAP # linear predictor (if the data followed the model expectations perfectly)
df$y <- round(df$exp_y + df$resids, digits = 0) # add the residuals

# check if equal to original data
df$Richness == df$y



# Quick way using R functions
df$preds <- predict(lme1)
df$y_from_preds <- round(df$preds + df$resids, digits = 0)
```







