---
title: "Homework 1: R Introduction"
author: "Answer Key"
output: pdf_document
---

*Adapted from University of Tennessee Biometry course by Dr. Jim Fordyce*

Turn in via blackboard next Friday by the end of the day. please title the document **hw1_yourlastname_yourfirstname**. Ideally it will be a PDF file generated from knitting an RMarkdown document.


## 1. Assign to a vector called X the values 1,3,5,6,7, and 9. Assign to a vector called Y the values 2,7,4,5,2 and 12.

```{r q1}

X <- c(1, 3, 5, 6, 7, 9)
Y <- c(2, 7, 4, 5, 2, 12)

```

 
## 2. Combine the two vectors together as one large vector called Z.

```{r q2}

Z <- c(X, Y)

```


## 3. Create a new vector that includes all the values in Z that are less than 6.

```{r q3}

Z_small <- Z[which(Z < 6)]

```


## 4. Using R functions, calculate the mean of Z.

```{r q4}

mean(Z)

```


## 5. Create an matrix where the columns are X and Y. Assign this matrix to an object called mat1.

```{r q5}

mat1 <- cbind(X, Y)

# alternate solution
mat1b <- matrix(c(X, Y), nrow = length(X), ncol = 2, byrow = FALSE)

```

## 6. Using the R function `rowSums`, calculate the sum of each row.

```{r q6}

rowSums(mat1)

```


## 7. Calculate the mean of the rows of mat1 and add this as a third column.

```{r q7}

means <- rowMeans(mat1)

mat1 <- cbind(mat1, means)

mat1

```


## 8. Extract the value of mat1 at the second row and first column.

```{r q8}

mat1[2, 1]

```


## 9. Write mat1 to a .csv file.

For my code to work you will have to have a folder called `Output` in your working directory.

```{r q9}

write.csv(mat1, file = "Output/hw1_mat.csv", row.names = FALSE)

```


## 10. Read the .csv file and assign it to an object called mat_new.

```{r q10}

mat_new <- read.csv(file = "Output/hw1_mat.csv")

```


## 11 Show that the third number in the first column of `mat_new` is equal to the fourth number in the second column.

```{r q11}

mat_new[3, 1] == mat_new[4, 2]

```

If this is not true, check that you set `row.names = FALSE` when writing out the object. Otherwise when reading it back in the first column will be the row numbers.

## 12 Make 'mat1' a data frame.

```{r q12}

mat1 <- as.data.frame(mat1)

```


## 13 Using R functions, calculate the length of X.

```{r q13}

length(X)

```


## 14 Assign the name "mean" to the third column of mat1.

```{r q14}

colnames(mat1)[3] <- "mean"

names(mat1)

```


## 15 Create a sequence of numbers from -10 to 10 with intevals of 0.5 using R functions. Assign it to an object and calculate the sum of the values.

```{r q15}

q15 <- seq(-10, 10, by = 0.5)

sum(q15)

```




