---
title: "Linear Mixed-Effects Models"
author: "Quantitative Analysis of Vertebrate Populations"
output:
  pdf_document: default
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Advantages of Mixed Models

- Increased scope of inference
- More honest accounting of uncertainty
- Efficiency of estimation

## Random Effects

- Random intercept
- Random slope(s) *not generally realistic - don't use*
- Random intercept and slope(s) independent
- Random intercept and slope(s) correlated

## Random intercept model

$$
y_i = \alpha_{j(i)} + \beta X_{i} + \epsilon_i
$$

where

$\alpha_j \sim \mathcal{N}(\mu_{\alpha}, \sigma_{\alpha}^2)$

$\epsilon_i \sim \mathcal{N}(0, \sigma^2)$

- snake $i$ at site $j$
- Mean size of the snake varies normally among populations

## Random intercept and slope model

$$
y_i = \alpha_{j(i)} + \beta X_{j(i)} + \epsilon_i
$$

where

$\alpha_j \sim \mathcal{N}(\mu_{\alpha}, \sigma_{\alpha}^2)$

$\beta_j \sim \mathcal{N}(\mu_{\beta}, \sigma_{\beta}^2)$

$\epsilon_i \sim \mathcal{N}(0, \sigma^2)$

## Simulate Data

```{r}
n.groups <- 56				# Number of populations
n.sample <- 10				# Number of vipers in each pop
n <- n.groups * n.sample 		# Total number of data points
pop <- gl(n = n.groups, k = n.sample) 	# Indicator for population
```

## Simulate Data

```{r}
# Body length (cm)
original.length <- runif(n, 45, 70) 
mn <- mean(original.length)
sd <- sd(original.length)
length <- (original.length - mn) / sd
hist(length, col = "grey")
```

## Set up design matrix

```{r}
Xmat <- model.matrix(~pop*length-1-length)
print(Xmat[1:5, 1:5], dig = 2) 		# Print 5 rows and columns
```

## Choose parameter values

```{r}
intercept.mean <- 230			# mu_alpha
intercept.sd <- 20			# sigma_alpha
slope.mean <- 60			# mu_beta
slope.sd <- 30				# sigma_beta
```

## Simulate effects

```{r}
intercept.effects<-rnorm(n = n.groups, 
                         mean = intercept.mean, 
                         sd = intercept.sd)
slope.effects <- rnorm(n = n.groups, mean = slope.mean, sd = slope.sd)
all.effects <- c(intercept.effects, slope.effects) # Put them all together
```

## Make linear model

```{r}
lin.pred <- Xmat[,] %*% all.effects	# Value of lin.predictor
eps <- rnorm(n = n, mean = 0, sd = 30)	# residuals 
mass <- lin.pred + eps			# response = lin.pred + residual
```

## Examine simulated data

```{r}
hist(mass, col = "grey")		# Inspect what we’ve created
```

## Examine simulated data

```{r}
library(lattice)
xyplot(mass ~ length | pop)
```


## Analysis under a random-intercepts model

```{r}
library(lme4)
lme.fit1 <- lmer(mass ~ length + (1 | pop), REML = TRUE)
lme.fit1
```


## Analysis under a random-coefficients model without correlation between intercept and slope

```{r}
lme.fit2 <- lmer(mass ~ length + (1 | pop) + ( 0+ length | pop))
lme.fit2
```

## Random-coefficients model with correlation between intercept and slope

$$
y_i = \alpha_{j(i)} + \beta X_{j(i)} + \epsilon_i
$$

where

$$
(\alpha_j, \beta_j) \sim MVN(\mu, \Sigma)
$$

$$
\mu = (\mu_{\alpha}, \mu_{\beta})
$$

$$
\left( \begin{array}{ccc}
\sigma_{\alpha}^2 & \sigma_{\alpha \beta} \\
\sigma_{\alpha \beta} & \sigma_{\beta}^2 \end{array} \right)
$$

$$
\epsilon_i \sim \mathcal{N}(0, \sigma^2)
$$

## Data generation

```{r}
n.groups <- 56
n.sample <- 10
n <- n.groups * n.sample 
pop <- gl(n = n.groups, k = n.sample)
```

## Data generation

```{r}
original.length <- runif(n, 45, 70) 	# Body length (cm)
mn <- mean(original.length)
sd <- sd(original.length)
cat("Mean and sd used to normalise.original length:", mn, sd, "\n\n")
length <- (original.length - mn) / sd
```

## Data generation - View distribution

```{r}
hist(length, col = "grey")
```

## Make model matrix

```{r}
library(MASS)				# Load MASS
# ?mvrnorm				# Check syntax
Xmat <- model.matrix(~pop*length-1-length)
# print(Xmat[1:21,], dig = 2) 		# Print top 21 rows
```


## Choose values

```{r}
intercept.mean <- 230			# Values for five hyperparameters
intercept.sd <- 20
slope.mean <- 60
slope.sd <- 30
intercept.slope.covariance <- 10
```

## Make correlated parameters

```{r}
mu.vector <- c(intercept.mean, slope.mean)
var.cova.matrix <- matrix(c(intercept.sd^2,intercept.slope.covariance, 
intercept.slope.covariance, slope.sd^2),2,2)

effects <- mvrnorm(n = n.groups, mu = mu.vector, Sigma = var.cova.matrix)
effects					# Look at what we’ve created
apply(effects, 2, mean)
var(effects)
```

## Make correlated parameters

```{r}
intercept.effects <- effects[,1]
slope.effects <- effects[,2]
all.effects <- c(intercept.effects, slope.effects) # Put them all together
```

## Make linear model

```{r}
lin.pred <- Xmat[,] %*% all.effects	# Value of lin.predictor
eps <- rnorm(n = n, mean = 0, sd = 30)	# residuals 
mass <- lin.pred + eps			# response = lin.pred + residual
```

## View histogram

```{r}
hist(mass, col = "grey")		# Inspect what we’ve created
```

## View lattice plot

```{r}
library("lattice")
xyplot(mass ~ length | pop)
```

## REML analysis using R

```{r}
lme.fit3 <- lmer(mass ~ length + (length | pop))
lme.fit3
```


## References

Kery, M. 2010. Introduction to WinBUGS for Ecologists. Academic Press.

Zuur et al. 2009. Mixed Effects Models and Extensions in Ecology with R. Springer.




