---
title: "Extracting Information from Regression Objects"
author: "Quantitative Analysis of Vertebrate Populations"
output:
  pdf_document: default
  ioslides_presentation:
    css: style.css
  beamer_presentation: default
  slidy_presentation: default
---


## Get the data

```{r}
sally <- read.table("Data/Salamander_Demographics.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",")
str(sally)
```

## Run a regression

Using the Salamander Demographic Data run a regression of the affect of svl on mass

```{r}
library(lme4)
lme1 <- lmer(mass ~ 1 + svl + (1 | plot), data = sally)
summary(lme1)
```

## Examine object and summary object

```{r}
str(lme1)
```


## Examine object and summary object

```{r}
lme1_sum <- summary(lme1)
lme1_sum
str(lme1_sum)
```


## Get Fitted and Residuals and Plot

```{r}
lme1_resids <- lme1_sum$residuals # or resid(lme1)
lme1_fits <- fitted(lme1)
par(mfrow = c(1, 2))
hist(lme1_resids)
plot(lme1_fits, lme1_resids)
abline(h = 0, col = "red")
par(mfrow = c(1,1))
```


## Get random effects

```{r}
lme1_ints <- ranef(lme1)
lme1_ints
mean(lme1_ints$plot[[1]]) # should approximate 0
```

## Get Coefficients

```{r}
coef(lme1)
fixef(lme1)
```

## Plot Random Effects

```{r}
Intercepts <- fixef(lme1)[1] + lme1_ints$plot[[1]]
Slope <- fixef(lme1)[2]
X <- seq(0, 60)
overall_intercept <- fixef(lme1)[1]
fit_line <- overall_intercept + Slope * X 
plot(X, fit_line, type = "l", lwd = 2, xlab = "SVL (mm)", ylab = "Mass (g)",
     xlim = c(30, 35),
     ylim = c(0.35, 0.6))
for(i in 1:length(Intercepts)) {
  rand_line <- Intercepts[i] + Slope * X
  lines(X, rand_line, col = "blue", lwd = 1)
}
lines(X, fit_line, lwd = 3)
```


## Exercise

* Simulate data with random slopes and intercepts
* Run a linear mixed model on the data
* Plot the expected lines for each group









