---
title: "AIC Practice Problems"
author: "Your Name Goes Here"
output:
  pdf_document: default
  html_document: default
---

## Data

These are salamander capture data from Great Smoky Mountains National Park along an elevational gradient. These are counts of *Eurycea wilderae*, the Blue-Ridge Two-Lined Salamander, on transects (25 m long by 2 m wide) and associated environmental and habitat conditions. Transects were in clusters by plot. You can use counts or the presence-absence data (counts coverted to 1 or 0 in the `presence` column).

```{r}
sal <- read.csv("Data/ewilderae.csv", stringsAsFactors = FALSE, header = TRUE)

str(sal)

unique(sal$plot)
length(unique(sal$plot)) # number of different plots
```

## Standardize the Continuous Independent Variables

Turn into a Z-score by substracting the mean from each observeration and dividing by the SD.

Now 95% of your observations will be between -2 and 2 (if your data were normally distributed)

$$
x^{\prime}_i = \frac{x_i - \bar x}{\sigma}
$$

```{r}
# check the range
range(sal$elev)
range(sal$temp)

# Use the scale function
# ?scale
sal$elev_s <- as.numeric(scale(sal$elev)) # without the as.numeric it creates a form that doesn't play well with others

# by hand
sal$temp_s <- (sal$temp - mean(sal$temp, na.rm = T)) / sd(sal$temp, na.rm = T) 

# make a custom function that standardizes the variable. 
# Requires a vector and outputs a standardized version of the vector
std <- function(var) {
  var_s <- (var - mean(var, na.rm = T)) / sd(var, na.rm = T)
  return(var_s)
}

# use custom function on a variable
sal$humidity_s <- std(sal$humidity)

# use custom function in combination with dplyr mutate function
library(dplyr)

sal <- sal %>%
  mutate(precip_s = std(precip),
         twi_s = std(twi),
         slope_s = std(slope))

# look at str and summary
str(sal)
summary(sal)

```

## Models

Transects within plots were part of the study design so plots should be included as random groups

```{r}
library(lme4)

glmer1 <- glmer(count ~ 1 + (1 | plot), data = sal, family = "poisson")
glmer2 <- glmer(count ~ 1 + temp_s + humidity_s + (1 | plot), data = sal, family = "poisson")
glmer3 <- glmer(count ~ 1 + elev_s + (1 | plot), data = sal, family = "poisson")
glmer4 <- glmer(count ~ 1 + elev_s + I(elev_s^2) + (1 | plot), data = sal, family = "poisson")
glmer5 <- glmer(count ~ 1 + slope_s + twi_s + (1 | plot), data = sal, family = "poisson")
glmer6 <- glmer(count ~ 1 + temp_s + humidity_s + elev_s + I(elev_s^2) + slope_s + twi_s + (1 | plot), data = sal, family = "poisson")
```

## AICc comparison

```{r}
library(AICcmodavg)

##set up candidate models          
Cand.mod <- list()

##global model          
Cand.mod[[1]] <- glmer1
Cand.mod[[2]] <- glmer2
Cand.mod[[3]] <- glmer3
Cand.mod[[4]] <- glmer4
Cand.mod[[5]] <- glmer5
Cand.mod[[6]] <- glmer6

##assign names to each model
Modnames <- c("intercept only", 
              "environ conditions",
              "elevation", 
              "quadratic elevation",
              "physical environ", 
              "global model") 

##model selection table based on AICc
aictab(cand.set = Cand.mod, modnames = Modnames)
```

## Check model fit

```{r}
plot(glmer4)
boxplot(resid(glmer4) ~ sal$plot,
        xlab = "Plot",
        ylab = "Pearson's Residuals")
```

## Random grouping needed?

Because the grouping was part of the experimental design, it is important to always include that in the analysis. However, if it wasn't part of the design but we thought their might be a reason to group certain plots we could still do the same thing as above. In that case, it might be worth seeing if it was worth the grouping.

Compare the best model from above with the same fixed effects model (same model just without any random effects).

```{r}
library(lme4)
glm4 <- glm(count ~ 1 + elev_s + I(elev_s^2), data = sal, family = "poisson")

##set up candidate models          
Cand.mod2 <- list()

##global model          
Cand.mod2[[1]] <- glmer4
Cand.mod2[[2]] <- glm4

##assign names to each model
Modnames2 <- c("GLMM", 
              "GLM") 

##model selection table based on AICc
# aictab(cand.set = Cand.mod2, modnames = Modnames2) # throws error

df <- data.frame(model = c("GLMM", "GLM"), AICc = c(AICc(glmer4), AICc(glm4)))
df[order(df$AICc), ]

```


## Grad Student Optional: Using Model Averaging for Multimodel Inference



