---
title: "Generalized Linear Mixed Models"
author: "Quantitative Analysis of Vertebrate Populations"
output:
  pdf_document: default
  ioslides_presentation:
    css: style.css
  beamer_presentation: default
  slidy_presentation: default
---

## Generalized Linear Model

Example with a Poisson distribution for count data. **Why do we use Poisson for counts?**

1. Distribution: $C_i \sim Poisson(\lambda_i)$
2. Link Function: $log(\lambda_i)$
3. Linear Predictor: $log(\lambda_i) = \alpha + \beta X_i$

## Generalized Linear Mixed Model

Add submodels for the random effect parameters.

1. Distribution: $C_i \sim Poisson(\lambda_i)$
2. Link Function: $log(\lambda_i)$
3. Linear Predictor: $log(\lambda_i) = \alpha_{j(i)} + \beta X_{i}$
4. Submodel for random effects parameters

$$
\alpha_j \sim \mathcal{N}(\mu_{\alpha}, \sigma_{\alpha}^2)
$$

## Random intercept and slope model

1. Distribution: $C_i \sim Poisson(\lambda_i)$
2. Link Function: $log(\lambda_i)$
3. Linear Predictor: $log(\lambda_i) = \alpha_{j(i)} + \beta_{j(i)} X_{i}$
4. Submodel for random effects parameters

$$
\alpha_j \sim \mathcal{N}(\mu_{\alpha}, \sigma_{\alpha}^2)
$$
$$
\beta_j \sim \mathcal{N}(\mu_{\beta}, \sigma_{\beta}^2)
$$


## Data generation (From Kery 2010)

Random slope and intercepts without correlation

```{r}
n.groups <- 16
n.years <- 30
n <- n.groups * n.years
pop <- gl(n = n.groups, k = n.years)
```

Standardize the year covariate to range from zero to one.

```{r}
original.year <- rep(1:n.years, n.groups)
year <- (original.year-1)/29
```

## Build design matrix without intercept

```{r}
Xmat <- model.matrix(~pop*year-1-year)
print(Xmat[1:21,], dig = 2) 		# Print top 21 rows
```

## Draw random intercept and slopes from normal distributions

```{r}
intercept.mean <- 3			# Choose values for the hyperparams
intercept.sd <- 1
slope.mean <- -2
slope.sd <- 0.6
intercept.effects<-rnorm(n = n.groups, mean = intercept.mean, sd = intercept.sd)
slope.effects <- rnorm(n = n.groups, mean = slope.mean, sd = slope.sd)
all.effects <- c(intercept.effects, slope.effects) # Put them all together
```

## Assemble and inspect results

```{r}
lin.pred <- Xmat[,] %*% all.effects	# Value of lin.predictor
C <- rpois(n = n, lambda = exp(lin.pred)) # Exponentiate and add Poisson noise
hist(C, col = "grey", xlab = "Counts", main = "")			# Inspect what we’ve created
```

## Assemble and inspect results

```{r}
library("lattice")
xyplot(C ~ original.year | pop, ylab = "Red-backed shrike counts", xlab = "Year")
```


## Analysis under a random-coefficients model

```{r, warning = FALSE, message=FALSE}
library('lme4')
glmm.fit <- glmer(C ~ year + (1 | pop) + ( 0+ year | pop), family = poisson)
summary(glmm.fit)				# Inspect results
```

## Examine Pearson residuals for patterning

```{r}
plot(glmm.fit)
```
