---
title: 'Generalized Linear Mixed Models: Binomial'
author: "Quantitative Analysis of Vertebrate Populations"
output:
  ioslides_presentation:
    css: style.css
  beamer_presentation: default
  pdf_document: default
  slidy_presentation: default
---

## Generalized Linear Model - Binomial

**=Logistic Regression**

* Remember the binomial distribution
* Series of "coin-flip" trials
* Count the total number of an outcome ("heads") in a series of $N$ trials ("coin flips")
* Estimate the Pr(heads) and factors that influence that
* Presence-Absence or detection-nondetection data

## Logistic Regression

1. Distribution: $C_i \sim Binomial(N, p_i)$
2. Link Function: $logit(p_i) = log\big(\frac{p_i}{1 - p_i}\big) = linear predictor$
3. Linear Predictor: $logit(p_i) = \alpha + \beta X_i$

Let's try it

## Data Generation

Two species (Cross-leaved gentian & Chiltern gentian) with P/A data from 50 sites

```{r}
N <- 50
p.cr <- 13/50 # avg prob of finding
p.ch <- 29/50 # avg prob of finding

C.cr <- rbinom(1, 50, prob = p.cr) # add binomial noise
C.ch <- rbinom(1, 50, prob = p.ch) # add binomial noise
C <- c(C.cr, C.ch)
species <- factor(c(0,1), labels = c("Cross-leaved", "Chiltern"))
```


## Analysis

Check if a significant difference in the distribution (freq of occurrence)

```{r}
summary(glm(cbind(C, N-C) ~ species, family = "binomial"))
predict(glm(cbind(C, N-C) ~ species, family = "binomial"), type = "response")
```


## Binomial GLMM

Random intercept and slope

1. Distribution: $C_i \sim Binomial(N_i, p_i)$
2. Link Function: $logit(p_i) = log\big(\frac{p_i}{1 - p_i}\big) = linear predictor$
3. Linear Predictor: $logit(p_i) = \alpha_{j(i)} + \beta_{j(i)} X_{i}$

$$
\alpha_j \sim \mathcal{N}(\mu_{\alpha}, \sigma_{\alpha}^2)
$$

$$
\beta_j \sim \mathcal{N}(\mu_{\beta}, \sigma_{\beta}^2)
$$

## Example: Generate Data

Red-backed Shrike: nest success (or failure) rather than counts

```{r}
n.groups <- 16
n.years <- 30
n <- n.groups * n.years
pop <- gl(n = n.groups, k = n.years)
```

Uniform covariate as an index of spring precipitation (0 = little rain, 1 = lots)

```{r}
precip <- runif(n, 0, 1)
```

Binomial total = number of nest attempts in survey year $i$

```{r}
N <- round(runif(n, 10, 50))
```

Build the design matrix

```{r}
Xmat <- model.matrix(~pop*precip-1-precip)
```

Choose parameter values and hyperparameters

```{r}
intercept.mean <- 1			# Select hyperparams
intercept.sd <- 1
slope.mean <- -2
slope.sd <- 1
intercept.effects<-rnorm(n = n.groups, mean = intercept.mean, sd = intercept.sd)
slope.effects <- rnorm(n = n.groups, mean = slope.mean, sd = slope.sd)
all.effects <- c(intercept.effects, slope.effects) # Put them all together
```

Assemble to make the linear predictor and visualize

```{r}
lin.pred <- Xmat %*% all.effects	# Value of lin.predictor
exp.p <- exp(lin.pred) / (1 + exp(lin.pred)) # Expected proportion

library("lattice")
xyplot(exp.p ~ precip | pop, ylab = "Expected woodchat shrike breeding success ",
xlab = "Spring precipipitation index", main = "Expected breeding success")
```

Add random binomial variation

```{r}
C <- rbinom(n = n, size = N, prob = exp.p) # Add binomial variation
xyplot(C/N ~ precip | pop, ylab = "Realized woodchat shrike breeding success ", 
xlab = "Spring precipipitation index", main = "Realized breeding success")
```


## Analysis under a random-coefficients model

Assume no correlation between intercept and slope

```{r}
library('lme4')
glmm.fit <- glmer(cbind(C, N-C) ~ precip + (1 | pop) + ( 0+ precip | pop), family = binomial)
summary(glmm.fit)
```


## Activity

Use the `DeerEcervi.txt` data and run a binomial GLMM including random intercepts by farm and fixed (non-random) independent predictor variables for Length, Sex, and Length x Sex.




